---
title: "Untitled"
format: revealjs
execute:
  echo: false
---

## Our Data{.smaller}

Read the _groceries.csv_ dataset.

Here you find the follwing variables:

- **turnover**: Weekly turnover (measured in 1000 NOK).
- **population**: Number of people living within 1 km of the store.
- **distance**: Distance (in km) to the nearest competing grocery store.
- **central**: Indicator variable equal to 1 if the store is located in a city center, 0 otherwise.
- **mall**: Indicator variable equal to 1 if the store lies within a mall (shopping center), 0
otherwise.

```{r}
library(car)
library(ggplot2)
library(GGally)
library(patchwork)
library(tidyverse)
rm(list = ls())
groceries = read.csv("groceries.csv")

```

Our interest lies in predicting the turnover based on the other available data.

## Exploratory analysis

- Create some plots to explore the relationship between turnover and the other variables
- Which quantities appear to influence the turnover?

##

```{r}
ggpairs(groceries)

```


## Simple linear regression

- Create a linear model where the only covariate is _distance_
- How do you interpret the two estimated parameters?
- One of the output of the model is the $R^2$ quantity, can you explain what it is?


## Simple linear regression{.smaller}

```{r}
model0 = lm(turnover ~ population, data = groceries)
summary(model0)

```

- If the population is 0, the turnover is ca 250 000 NOK per week (??)
- For each extra person leaving within 1 Km from the shop, the turnover increases by ca 390 NOK

## Simple linear regression

- Plot the data together with the fitted regression line
- Add prediction intervals to the line, what do you learn?
- What is the predicted turnover (with uncertainty) for a shop that in an area where 2000 people live?


## Simple linear regression

```{r}

# Create values of population over the range of your data
newdata <- data.frame(
  population = seq(
    min(groceries$population),
    max(groceries$population),
    length.out = 100
  )
)

# Get fitted values and prediction intervals
pred <- predict(model0, newdata = newdata, interval = "prediction")

# Add predictions to newdata
newdata <- cbind(newdata, as.data.frame(pred))

# Plot
ggplot() +
  geom_point(data = groceries, aes(x = population, y = turnover)) +
  geom_ribbon(
    data = newdata,
    aes(x = population, ymin = lwr, ymax = upr),
    alpha = 0.2
  ) +
  geom_line(
    data = newdata,
    aes(x = population, y = fit),
    linewidth = 1
  ) +
  labs(
    x = "Population",
    y = "Turnover"
  ) +
  theme_minimal()
```

```{r}

pred <- predict(model0, newdata = data.frame(population = 2000), interval = "prediction")
pred
```

## Simple linear regression  - residuals

The residuals, in this case,  are defined as the observed minus the predicted turnover

- Compute the residuals and plot them
- Create a QQ plot of the residuals, what does this indicate?
- What are we looking for in a residual plot?
- Which model assumption are we checking when looking at residuals?


## Simple linear regression  - residuals
```{r}
res = data.frame(residuals = residuals(model0), 
                 pred = predict(newdata = groceries, model0),
                 obs = groceries$turnover,
                 pop = groceries$population,
                 ii = 1:dim(groceries)[1])

p1 = res %>% ggplot() + geom_point(aes(ii, residuals)) + ggtitle("Residuals")
p2 = res %>% ggplot() + geom_point(aes(pred, residuals)) + ggtitle("Residuals vs Predict")


p3 <- ggplot(res, aes(sample = residuals))
p3 = p3 + stat_qq() + stat_qq_line()+xlab("Theoretical Quantiles") + ylab("Sample Quantiles") +
  ggtitle("QQ plot")
 p1+p2 + p3 + plot_layout(ncol= 2)

```

## Model assumption and residuals{.smaller}

Remember:
$$
y_i = \underbrace{\beta_0 + \beta_1\ x_i}_{\text{Sistematic part}} + \underbrace{\epsilon_i}_{\text{Random part/ error}}, \qquad \epsilon_i\sim\mathcal{N}(0,\sigma^2)
$$

The key assumption we check with residuals is:

**The errors should behave like (gaussian) random noise.**

A residual plot should therefore show:

- Random scatter around 0
- Roughly the same amount of spread across the range
- No obvious curve or other pattern
- No "funnel" shape


## Improving the model

Add the _distance_ covariate to the model

- How do you interpret the estimated parameters?
- What happened to the parameter for population? Why?
- Check if this has improved the model using $R^2$

- You could consider instead to use _log(distance)_, why would that make sense?
- Plot the predicted turnover for both model when population is 1000 and the distance goes from 0 to 10 km. (You can add prediction uncertainties). What do you notice?

## Improving the model


```{r}
model1 = lm(turnover ~ population + distance, data = groceries)
model1b = lm(turnover ~ population + log(distance), data = groceries)
```


Model 0 (only population)


```{r}
summary(model0)
```


Model 1 ( population + distance)


```{r}
summary(model1)
```

## Improving the model


```{r}
newdata = data.frame(distance  = seq(0.2, 10, 0.01), population = 1000)
pred1 = data.frame(predict(model1, newdata, interval = "pred" )) %>%
  mutate(distance = newdata$distance)
pred2 = data.frame(predict(model1b, newdata, interval = "pred")) %>%
  mutate(distance = newdata$distance)


ggplot() + geom_ribbon(data = pred1, aes(x = distance, ymin = lwr, ymax = upr), alpha = 0.5)+ 
  geom_line(data = pred1, aes(x = distance,y = fit)) + 
  geom_ribbon(data = pred2, aes(x = distance, ymin = lwr, ymax = upr), alpha = 0.5, fill = "red") +
  geom_line(data = pred2, aes(x = distance,y = fit), color ="red") 



```

## Adding categorical covariates

We want to check weather being in a mall or being in the city center has also an effect on the turnover.

- Fit a model where you have population, log(distance) and you add the variable _mall_
- How do you interpred the pararameter for mall?


## Adding categorical covariates

```{r}
model2 = lm(turnover ~ population + log(distance) + mall, data = groceries)
summary(model2)
```